---
title: "Synchronous and asynchronous model runs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Synchronous and asynchronous model runs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE          # all chunks need a live server + token; shown, not run
)
```

```{r setup}
library(modelscloud)
```

## Two ways to run a model

Some ModelsCloud models can be executed in either of two modes:

- **Synchronous** (the default): you call `model_run()` and R **blocks** until
  the server finishes, then hands you the result. Simple, and ideal for quick
  models.
- **Asynchronous**: `model_run(async = TRUE)` **returns immediately** with a
  lightweight *job handle*. The model keeps running on the server while your R
  session is free to do other things. You fetch the result later with
  `get_async_results()`.

Async exists for long-running models — policy simulations, large agent counts,
anything that might take minutes. Blocking an interactive session (or hitting an
HTTP timeout) on such a run is painful; submitting it as a job is not.

Throughout this vignette the example is `epicrpexa`, a COPD policy
microsimulation that supports both modes.

```{r connect}
connect_to_model("mohsenss/epicrpexa", access_key = "YOUR_API_KEY")
```

(You can also store your key once in `.Renviron` as `MODELSCLOUD_ACCESS_KEY`
and omit `access_key` — see the README.)

## Synchronous: the baseline

Nothing special is required. `model_run()` waits and returns the decoded result:

```{r sync}
input  <- get_default_input()
result <- model_run(input)        # blocks until the server is done

result$basic$total_qaly
```

This is fine when the model returns in a few seconds. For a small agent count it
will. For a realistic policy run it won't — which is where async comes in.

## Asynchronous: submit, then retrieve

Set `async = TRUE`. `model_run()` returns a **job handle** right away:

```{r async-submit}
job <- model_run(input, async = TRUE)
#> Async job submitted (executionId: 6a2a...). Retrieve it with get_async_results().

class(job)
#> [1] "pexa_result_async" "pexa_result"
```

`job` is *not* your result — it is a receipt. Pass it to `get_async_results()`
to fetch the actual output once the server has finished.

### Option A — block until it's done (`wait = TRUE`)

The simplest pattern. `get_async_results()` polls the server for you and returns
only when the job is complete:

```{r async-wait}
result <- get_async_results(job, wait = TRUE)

result$basic$total_qaly
```

You can do the whole thing in one line:

```{r async-oneline}
result <- get_async_results(model_run(input, async = TRUE), wait = TRUE)
```

`wait = TRUE` polls every `interval` seconds (default 2) up to `timeout` seconds
(default 300). If the timeout is reached the job handle is returned with a
warning, so you can keep polling rather than losing the job:

```{r async-timeout}
# Wait up to 30 minutes, checking every 10 seconds
result <- get_async_results(job, wait = TRUE, interval = 10, timeout = 1800)
```

### Option B — check once, poll manually

By default (`wait = FALSE`) `get_async_results()` checks **once** and returns
immediately:

- if the job has finished, you get the decoded result;
- if it is still running, you get the job handle back, unchanged.

This lets you do other work between checks, or drive your own polling loop:

```{r async-manual}
result <- get_async_results(job)            # one check

while (inherits(result, "pexa_result_async")) {
  message("still running...")
  Sys.sleep(5)
  result <- get_async_results(job)          # check again
}

result$basic$total_qaly                     # finished
```

The test `inherits(result, "pexa_result_async")` is how you tell "still a job
handle" from "this is the finished result".

## Choosing the mode per call vs. for the session

`async` is decided **per call** on `model_run()`. You do not need to reconnect
to switch modes:

```{r per-call}
quick <- model_run(small_input)                  # sync
job   <- model_run(big_input, async = TRUE)      # async, same session
```

If most of your runs are long, set async as the **session default** with
`connect_to_model()`, and a bare `model_run()` becomes asynchronous:

```{r session-default}
connect_to_model("mohsenss/epicrpexa", access_key = "YOUR_API_KEY",
                 async = TRUE)

job <- model_run(input)             # async, because that's the session default
```

`connect_to_model(async = ...)` only sets the *default*. An explicit argument on
`model_run()` always wins, so you can still force a single call back to sync:

```{r override}
result <- model_run(input, async = FALSE)   # sync, despite the async default
```

The resolution order is: **`model_run()`'s `async` argument → session default
from `connect_to_model()` → `FALSE`**.

## A realistic large run

The reason to reach for async — a high agent count that takes a while. Set the
agent count on the input, submit asynchronously, and block until done:

```{r big-run}
input <- get_default_input()
input$n_agents <- 1e6          # a million agents

result <- get_async_results(
  model_run(input, async = TRUE),
  wait = TRUE, timeout = 3600  # allow up to an hour
)

result$basic$total_qaly
```

## Plots work the same way

If a model produces plots, [get_plots()] works on an async result exactly as it
does on a synchronous one — the server response needed to fetch plots travels
with the decoded result:

```{r plots}
result <- get_async_results(model_run(input, async = TRUE), wait = TRUE)

get_plots(result)            # list available plots
get_plots(result, id = 1)    # fetch and display one
```

## Cheat sheet

| Task | Call |
|---|---|
| Run and wait | `model_run(x)` |
| Submit a job | `job <- model_run(x, async = TRUE)` |
| Block until done | `get_async_results(job, wait = TRUE)` |
| Check once (non-blocking) | `get_async_results(job)` |
| Is it still running? | `inherits(res, "pexa_result_async")` |
| Make async the session default | `connect_to_model(..., async = TRUE)` |
| Force one call back to sync | `model_run(x, async = FALSE)` |

## See also

- `?model_run`
- `?get_async_results`
- `?connect_to_model`
