---
title: "Example: EPIC-R COPD policy model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example: EPIC-R COPD policy model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE          # needs a live server + a key for the mohsenss namespace
)
```

This vignette shows `modelscloud` against a **real** policy model:
[EPIC](https://resp.core.ubc.ca/research/EPIC), a whole-disease microsimulation
of chronic obstructive pulmonary disease (COPD). It is served as
`mohsenss/epicrpexa`, a wrapper around the `epicR` package.

Where QRISK3 (see the other example vignette) is a quick prediction model, EPIC
is a simulation whose runtime grows with the number of simulated agents — so it
is a natural fit for **asynchronous** execution.

This model is outside the public `examples` collection, so you'll need your own
API key.

```{r setup}
library(modelscloud)

connect_to_model("mohsenss/epicrpexa", access_key = "YOUR_API_KEY")
```

## Default input

Policy models don't take a patient dataset; they take a set of parameters.
`get_default_input()` returns a ready-to-use `model_input` — a named list whose
`input` element holds EPIC's default parameter set:

```{r default}
mi <- get_default_input()
names(mi)
#> [1] "input"

mi$input$global_parameters$time_horizon
#> [1] 20
```

You adjust this list and pass it to `model_run()`. Scalar controls of the
underlying simulation — such as `n_agents`, `seed`, `time_horizon` — are set as
top-level elements of `model_input`; the server expands them onto the simulator.

## A light run (synchronous)

With a modest agent count the model returns quickly, so a plain call works:

```{r sync}
mi <- get_default_input()
mi$n_agents <- 1000

result <- model_run(mi)
result$basic$total_qaly
```

`model_run()` returns a list with a `$basic` summary (and `$extended` if you
request it by setting `mi$extended_results <- TRUE`).

## A large run (asynchronous)

For a realistic policy run — say ten million agents — submit the job with
`async = TRUE` and retrieve it when it finishes:

```{r async}
mi <- get_default_input()
mi$n_agents <- 1e7

job    <- model_run(mi, async = TRUE)            # returns immediately
result <- get_async_results(job, wait = TRUE,    # block & poll until done
                            timeout = 3600)
result$basic$total_qaly
```

To **check progress** without blocking, call `get_async_results(job)` on its
own: it returns the finished result if the run is done, or the job handle
unchanged if it is still running.

```{r progress}
res <- get_async_results(job)
if (inherits(res, "pexa_result_async")) message("still running")
```

See the *Synchronous and asynchronous model runs* vignette for the full async
workflow (manual polling loops, timeouts, session defaults).
