---
title: "Getting started with modelscloud"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with modelscloud}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE          # all chunks need a live server; shown, not run
)
```

`modelscloud` lets you call models hosted on the ModelsCloud platform as if they
were local R functions. This vignette walks through the basics using the public
**`examples`** collection — two minimal toy models you can run with a shared
public test key, no signup required.

```{r setup}
library(modelscloud)

# Public test key for the `examples` collection
KEY <- "23b7bab3-118e-4516-b53c-91bca8e0082d"
```

## 1. Connect to a model

`connect_to_model()` stores the model path and your key for the rest of the
session, so you don't repeat them on every call. It's optional (you can pass
`model_path` / `access_key` to each function instead) but it keeps scripts tidy.

```{r connect}
connect_to_model("examples/toymodel1", access_key = KEY)
```

Model paths are `"<collection>/<model>"`, e.g. `"examples/toymodel1"`.

## 2. A prediction model (synchronous)

`toymodel1` predicts a risk for one or more patients from `sex`, `age`, and
`marker_value`. Two functions cover the workflow:

- `get_sample_input()` — an example input you can run as-is
- `model_run()` — runs the model and returns the result

```{r pred}
sample <- get_sample_input()
sample
#>   sex age marker_value
#> 1   0  55          1.2
#> 2   1  62          2.4
#> 3   1  48          0.7

result <- model_run(sample)   # model_input is the first argument
result
#>   sex age marker_value    risk
#> 1   0  55          1.2 0.05787
#> 2   1  62          2.4 0.27289
#> 3   1  48          0.7 0.04565
```

The result comes back as a native R data frame — class information is preserved
across the network, so you can work with it exactly as if the model had run
locally.

## 3. Retrieving plots a model produced

Some models draw plots while they run. `toymodel1`, for instance, produces a
barplot of the predicted risks. OpenCPU captures any such plot on the server,
and you fetch it with `get_plots()` — which *retrieves* the plot but does not
draw it. To display it, call `plot()` on the retrieved image:

```{r plots}
result <- model_run(get_sample_input())

get_plots(result)                  # list the plots this run produced
img <- get_plots(result, id = 1)   # retrieve the first one
plot(img)                          # display it
```

Calling `get_plots(result)` with no `id` lists what's available; passing an
`id` returns that plot's image, which you then `plot()`.

## 4. A policy model with default inputs

`toymodel2` is a small policy/economic simulation. Instead of a sample dataset
it exposes `get_default_input()` — a ready-to-use set of parameters you can
tweak before running.

```{r policy}
connect_to_model("examples/toymodel2", access_key = KEY)

input <- get_default_input()
names(input)
#> "n_agents" "time_horizon" "p_event" "cost_event" ...

input$n_agents <- 5000        # adjust any parameter
result <- model_run(input)
result$total_qaly
```

## 5. Supplying credentials without connecting

If you prefer not to call `connect_to_model()`, pass `model_path` and
`access_key` directly. An explicit argument always overrides the stored value.

```{r explicit}
sample <- get_sample_input(model_path = "examples/toymodel1", access_key = KEY)
result <- model_run(sample, model_path = "examples/toymodel1", access_key = KEY)
```

You can also store your key once in `.Renviron` as `MODELSCLOUD_ACCESS_KEY`;
`modelscloud` will pick it up automatically, so a bare
`connect_to_model("examples/toymodel1")` is enough.

## Where next

- **Synchronous and asynchronous model runs** — for models that may take a
  while, submit a job and poll for the result.
- **Example: QRISK3** and **Example: EPIC-R** — real deployed models.

```{r more, eval = FALSE}
vignette("async", package = "modelscloud")
```
