---
title: "Example: QRISK3 cardiovascular risk model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example: QRISK3 cardiovascular risk model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE          # needs a live server + a key for the mohsenss namespace
)
```

This vignette shows `modelscloud` against a **real** deployed model:
[QRISK3](https://qrisk.org/), which estimates an individual's 10-year risk of
developing cardiovascular disease. It is served as `mohsenss/qrisk3pexa`, a
wrapper around the [`QRISK3`](https://cran.r-project.org/package=QRISK3) R
package.

Unlike the toy models in the *Getting started* vignette, this model lives
outside the public `examples` collection, so you'll need your own API key.

```{r setup}
library(modelscloud)

connect_to_model("mohsenss/qrisk3pexa", access_key = "YOUR_API_KEY")
```

## Sample input

QRISK3 takes a table of patients. `get_sample_input()` returns a ready-to-use
example dataset (the `QRISK3` package's bundled test patients, with column names
matching the model's API):

```{r sample}
patients <- get_sample_input()
head(patients)
#>   patid gender age atrial_fibrillation weight height ethiniciy ...
#> 1     1      1  64                   0     80    178         1 ...
#> 2     2      0  58                   1     72    165         1 ...

nrow(patients)
#> [1] 19
```

Pass `n` to get just the first few rows:

```{r sample-n}
get_sample_input(n = 3)    # first 3 patients
```

## Running the model

`model_run()` predicts every patient in the input and returns a data frame with
the QRISK3 score columns added:

```{r run}
results <- model_run(patients)
head(results[, c("ID", "QRISK3_2017")])
#>   ID QRISK3_2017
#> 1  1    17.22985
#> 2  2    36.01234
#> 3  3     9.84512
```

Because the result is a native R data frame, you can summarise or plot it
locally as usual:

```{r summary}
summary(results$QRISK3_2017)
hist(results$QRISK3_2017,
     main = "Predicted 10-year CVD risk", xlab = "QRISK3 score (%)")
```

## One-liner

`get_sample_input()` feeds straight into `model_run()`:

```{r oneliner}
results <- model_run(get_sample_input())
```

QRISK3 is a prediction model: every call is **synchronous** — `model_run()`
returns the predictions directly. For a model that supports asynchronous
execution, see the *Example: EPIC-R* vignette.
